An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-10-29 VLDB2026/main.tex
change bib
ok

\usepackage{multirow}
% \usepackage{svg} % For SVG support
\usepackage{graphicx}
\usepackage{ulem}  % 导言区加上这一行
\usepackage{microtype}  %处理换行
\usepackage{subcaption} % 需要subcaption宏包支持子图
\usepackage{caption}
\usepackage{fontawesome}
\usepackage{makecell}
\usepackage{array}

@@

	\author{Yan Li}
	\affiliation{%
		\institution{Huazhong University of Science and Technology}
	}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.2cm}
		\setlength{\belowcaptionskip}{-0.1cm}

@@

		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.3cm}		
		\caption{Graph Index and IVF Index}

@@

		\centering
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}

@@

	\noindent\textbf{\underline{NHQ} \footnote{\textcolor{black}{https://github.com/KGLab-HDU/TKDE-under-review-Native-Hybrid-Queries-via-ANNS}} \cite{NHQ}}
	Traditional attribute filtering algorithms usually perform attribute constraints and ANN search separately. In contrast, NHQ is the first to implement simultaneous filtering, integrating both aspects into a unified framework. NHQ constructs an index based on a nearest neighbor graph and introduces a fusion distance that jointly captures vector similarity and attribute similarity. Leveraging this fusion distance, NHQ unifies vector similarity and attribute matching into a single comprehensive similarity measure and builds the graph accordingly. During the query process, NHQ efficiently prunes irrelevant edges via this composite index, enabling fast retrieval of results that satisfy both vector similarity and attribute constraints.

@@

\noindent\textbf{\underline{Faiss}\footnote{\textcolor{black}{https://github.com/facebookresearch/faiss}} \cite{Faiss}}
Faiss is an efficient similarity search library. It supports hybrid search by using an ID selector to filter data before performing ANN search.
In Faiss, we use the HQI batch optimization strategy in the IVF index. It groups queries with the same filter conditions. Each group only runs the filter once. It uses efficient matrix operations to compute vector similarity in batches.

\noindent\textbf{\underline{PASE} \footnote{\textcolor{black}{https://github.com/alipay/PASE}} \cite{pase}} PASE is a vector index plugin for PostgreSQL. It uses a post-filtering strategy. Since the candidate set size is fixed, it may fail to return enough top-k results.

\noindent\textbf{\underline{VBASE} \footnote{\textcolor{black}{https://github.com/microsoft/MSVBASE}} \cite{vbase}} VBASE is also a vector plugin for PostgreSQL. Unlike PASE, it applies filters during the HNSW index traversal, discarding unmatched nodes on the fly. This strategy ensures that enough valid results are returned .

\noindent\textbf{\underline{Milvus} \footnote{\textcolor{black}{https://github.com/milvus-io/milvus}} \cite{milvus}} Milvus is a vector database designed for large-scale similarity search. In hybrid queries, it adopts a pre-filtering strategy, similar to the ID selector mechanism used in Faiss.
%It filters data first, then performs approximate nearest neighbor search. This supports hybrid search.

%	The vector libraries and databases discussed in the following are not explicitly designed for hybrid queries, but they support both attribute filtering and range filtering. Additionally, all the functionalities shown in Table 1 are also supported.
%	
%	\noindent\textbf{\underline{Faiss} \footnote{\textcolor{black}{https://github.com/facebookresearch/faiss}} \cite{Faiss}}
%	Faiss is a library designed for efficient similarity search. It supports various indexing structures, including IVF, Product Quantization (PQ) \cite{PQ}, and HNSW. In hybrid query scenarios, Faiss supports vector search with an ID selector, which is a bitmap aligned with the dataset size. Users can generate this selector by first applying custom attribute filtering methods. During the query process, Faiss excludes vectors based on the selector, enabling efficient integration of attribute filtering with ANN search. In addition, we adopt the batch optimization strategy of HQI~\cite{HQI} for the IVF index in Faiss. This strategy groups queries with the same filter conditions, performs filtering once per group, and then uses efficient matrix operations to perform batch vector similarity calculations.
%	
%	
%	\noindent\textbf{\underline{PASE} \footnote{\textcolor{black}{https://github.com/alipay/PASE}} \cite{pase}}
%	PASE is a vector indexing plugin for the PostgreSQL database \cite{postgresql13.4}, supporting two index types: IVF\_Flat \cite{johnson2019billion} and HNSW. By leveraging database capabilities, PASE enables both attribute filtering and range filtering. We focus on the post-filtering strategy based on HNSW, which facilitate comparison with other methods. When a query request is received, PASE first obtains a candidate set using the HNSW index. It then applies the filtering conditions from the WHERE clause to refine the results. However, since the candidate set has a fixed size, low-selectivity filters may result in fewer than $k$ results being returned.
%	
%	
%	\noindent\textbf{\underline{VBASE} \footnote{\textcolor{black}{https://github.com/microsoft/MSVBASE}} \cite{vbase}}
%	Similar to PASE, VBASE is a PostgreSQL-based vector indexing plugin supporting HNSW, SPTAG~\cite{sptag}, and SPANN~\cite{spann}. We focus on HNSW-based search because SPTAG and SPANN have poorer performance and their source code does not fully implement index-building functionality. VBASE adopts a post-filtering strategy. However, it employs an iterative filtering mechanism: during index traversal, each node is checked against the WHERE clause, and non-matching nodes are discarded. This avoids the issue in PASE where the final result set may contain fewer than $k$ results.
%	
%	\noindent\textbf{\underline{Milvus} \footnote{\textcolor{black}{https://github.com/milvus-io/milvus}} \cite{milvus}}
%	Unlike PASE and VBASE, which are database extensions, Milvus is an vector database purpose-built for large-scale similarity search. It supports multiple index types, including HNSW, IVF\_Flat, and IVF\_PQ, and provides extensive optimization for real-world deployment. Among its indexes, IVF\_Flat is commonly used due to its balance between performance and simplicity. In hybrid queries, Milvus first pre-filters the dataset using user-specified conditions, then traverses only the corresponding clusters for ANN search—similar to the ID selector mechanism of Faiss.

@@

		%		\caption{Time and space overhead of attribute filtering index construction}
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		\caption{\textcolor{black}{Time and Space Overhead (AF)}}
		\label{fig:build_index_comparison}
	\end{figure}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Performance of Single-Attribute Index Construction and Single-Attribute Query }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

@@

	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.3cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.85\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_4_1_MultiLabel_1thread.pdf}
		\caption{Performance of Multi-Attribute Index Construction and Multi-Attribute Query}
		\label{fig:exp_4_1_MultiLabel_1thread}
	\end{figure*}

@@

		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.15cm}

@@

\begin{figure*}
	\centering
	\setlength{\abovecaptionskip}{0.0cm}
	\begin{minipage}[t]{0.37\textwidth}
		\centering
		\setlength{\abovecaptionskip}{-0.2cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\caption{Multi-Modal Datasets (AF)}
		\label{fig:attribute-multimodal}  
	\end{minipage}%
	\hfill
	\begin{minipage}[t]{0.27\textwidth}
		\centering
		\setlength{\abovecaptionskip}{-0.2cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\caption{Large-Scale Dataset (AF)}
		\label{fig:attribute big dataset}
	\end{minipage}
	\hfill 
	\begin{minipage}[t]{0.35\textwidth}
		\centering
		\setlength{\abovecaptionskip}{-0.2cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\caption{Different Platforms (AF)}
		\label{fig:attribute-cross-platform}
	\end{minipage}%
\end{figure*}

@@

\footnotesize{
	\begin{minipage}{\linewidth}
		\vspace{0.1cm}
		\textit{Performance Levels:} \textbf{B} = Best, G = Good , M = Moderate, P = Poor, - = Not Supported;
		\textit{Performance-Critical}: Requires low latency and high recall;
		\textit{Index-Efficient}: Emphasizes lightweight index construction;
		\textit{Memory-Constrained}: Operates under limited memory conditions;
		\textit{Large-Scale}: For scenarios involving datasets at the scale of 100 million vectors;
		\textit{Multi-Modal}: For multi-modal dataset scenarios;
%		\textit{High-LID}: For complex dataset scenarios;
		\textit{Rare Attribute / Low Range}: For filtering with low attribute selectivity or small ranges;
		\textit{Distribution-Robust}: For scenarios with variable attribute distributions;
		\textit{General-Purpose}: Balanced performance across varied scenarios.

>>>>

\usepackage{multirow}
% \usepackage{svg} % For SVG support
\usepackage{graphicx}
\usepackage{ulem}  % 导言区加上这一行
\usepackage{microtype}  %处理换行
\usepackage{subcaption} % 需要subcaption宏包支持子图
\usepackage{caption}
\usepackage{fontawesome}
\usepackage{makecell}
\usepackage{array}
\usepackage[all]{nowidow}

@@

	\author{Yan Li}
	\affiliation{%
		\institution{Wuhan University of Technology}
	}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.15cm}
		\setlength{\belowcaptionskip}{-0.10cm}

@@

	\setlength{\abovecaptionskip}{0.15cm}
	\setlength{\belowcaptionskip}{-0.10cm}
		\caption{Graph Index and IVF Index}

@@

		\centering
		\footnotesize{
			\begin{minipage}{\linewidth}

@@

	\noindent\textbf{\underline{NHQ} \footnote{https://github.com/KGLab-HDU/TKDE-under-review-Native-Hybrid-Queries-via-ANNS} \cite{NHQ}}
	Traditional attribute filtering algorithms usually perform attribute constraints and ANN search separately. In contrast, NHQ is the first to implement simultaneous filtering, integrating both aspects into a unified framework. NHQ constructs an index based on a nearest neighbor graph and introduces a fusion distance that jointly captures vector similarity and attribute similarity. Leveraging this fusion distance, NHQ unifies vector similarity and attribute matching into a single comprehensive similarity measure and builds the graph accordingly. During the query process, NHQ efficiently prunes irrelevant edges via this composite index, enabling fast retrieval of results that satisfy both vector similarity and attribute constraints.

@@

\noindent\textbf{\underline{Faiss}\footnote{\url{https://github.com/facebookresearch/faiss}} \cite{Faiss}}
Faiss is an efficient similarity search library. It supports hybrid search by using an ID selector to filter data before performing ANN search.
In Faiss, we use the HQI batch optimization strategy in the IVF index. It groups queries with the same filter conditions. Each group only runs the filter once. It uses efficient matrix operations to compute vector similarity in batches.

\noindent\textbf{\underline{PASE}\footnote{\url{https://github.com/alipay/PASE}} \cite{pase}}  PASE is a vector index plugin for PostgreSQL. It uses a post-filtering strategy. Since the candidate set size is fixed, it may fail to return enough top-k results.

\noindent\textbf{\underline{VBASE} \footnote{\url{https://github.com/microsoft/MSVBASE}} \cite{vbase}} VBASE is also a vector plugin for PostgreSQL. Unlike PASE, it applies filters during the HNSW index traversal, discarding unmatched nodes on the fly. This strategy ensures that enough valid results are returned .

\noindent\textbf{\underline{Milvus} \footnote{\url{https://github.com/milvus-io/milvus}} \cite{milvus}} Milvus is a vector database designed for large-scale similarity search. In hybrid queries, it adopts a pre-filtering strategy, similar to the ID selector mechanism used in Faiss.
%It filters data first, then performs approximate nearest neighbor search. This supports hybrid search.

@@

		%		\caption{Time and space overhead of attribute filtering index construction}
		\setlength{\abovecaptionskip}{0.15cm}
		\setlength{\belowcaptionskip}{-0.10cm}
		\caption{\textcolor{black}{Time and Space Overhead (AF)}}
		\label{fig:build_index_comparison}
	\end{figure}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.15cm}
		\setlength{\belowcaptionskip}{-0.10cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Performance of Single-Attribute Index Construction and Single-Attribute Query }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

@@

	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.15cm}
		\setlength{\belowcaptionskip}{-0.10cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.85\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.15cm}
		\setlength{\belowcaptionskip}{-0.10cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_4_1_MultiLabel_1thread.pdf}
		\caption{Performance of Multi-Attribute Index Construction and Multi-Attribute Query}
		\label{fig:exp_4_1_MultiLabel_1thread}
	\end{figure*}

@@

		\setlength{\abovecaptionskip}{0.15cm}
		\setlength{\belowcaptionskip}{-0.10cm}

@@

\begin{figure*}
	\centering
%	\setlength{\abovecaptionskip}{0.0cm}
	\begin{minipage}[t]{0.37\textwidth}
		\centering
		\setlength{\abovecaptionskip}{-0.2cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\caption{Multi-Modal Datasets (AF)}
		\label{fig:attribute-multimodal}  
	\end{minipage}%
	\hfill
	\begin{minipage}[t]{0.27\textwidth}
		\centering
		\setlength{\abovecaptionskip}{-0.2cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\caption{Large-Scale Dataset (AF)}
		\label{fig:attribute big dataset}
	\end{minipage}
	\hfill 
	\begin{minipage}[t]{0.35\textwidth}
		\centering
		\setlength{\abovecaptionskip}{-0.2cm}
		\setlength{\belowcaptionskip}{-0.25cm}
		\caption{Different Platforms (AF)}
		\label{fig:attribute-cross-platform}
	\end{minipage}%
\end{figure*}

@@

\footnotesize{
	\begin{minipage}{\linewidth}

		\textit{Performance Levels:} \textbf{B} = Best, G = Good , M = Moderate, P = Poor, - = Not Supported;
		\textit{Performance-Critical}: Requires low latency and high recall;
		\textit{Index-Efficient}: Emphasizes lightweight index construction;
		\textit{Memory-Constrained}: Operates under limited memory conditions;
		\textit{Large-Scale}: For scenarios involving datasets at the scale of 100 million vectors;
		\textit{Multi-Modal}: For multi-modal dataset scenarios;
%		\textit{High-LID}: For complex dataset scenarios;
		\textit{Rare Attribute / Low Range}: For filtering with low attribute selectivity or small ranges;
		\textit{Distribution-Robust}: For scenarios with variable attribute distributions;
		\textit{General-Purpose}: Balanced performance across varied scenarios.
